### 2022-12-20 Manuscript/Example_from_Chris.Rmd
updating plotting code and manuscript

---
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
output:
  bookdown::pdf_document2:
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
  bibliography: IBMWG.bib
  word_document: default
  html_document: default
  pdf_document:
    toc: no
fontsize: 12pt
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage[belowskip=0pt,aboveskip=0pt]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}

#read in results tables
YT_results<-read.csv("Yellowtail_results.csv")
Cod_results<-read.csv("Cod_results.csv")
Had_results<-read.csv("Haddock_results.csv")
```

@@

# Estimating Population Trends with Stratified Random Sampling Under the Pressures of Climate Change

Benjamin A. Levy^1^, Christopher M. Legault^2^, Timothy J. Miller^2^, Elizabeth N. Brooks^2^

^1^Ben's Institution, USA  
^2^National Marine Fisheries Service, Northeast Fisheries Science Center, Woods Hole, MA, USA

Corresponding author: Ben Levy (benjamin.levy@noaa.gov)

@@

<!-- Table1 <- YT_results -->

<!-- write.table( -->
<!--    format(f, digits=2), "", -->
<!--    sep="|", row.names=F, col.names=F, quote=F, eol="|\n") -->

@@

An Abstract

@@

keyword 1, keyword 2

@@

```{r}
knitr::kable(YT_results, caption = "Yellowtail error results")

knitr::kable(Cod_results, caption = "Cod error results")

knitr::kable(Had_results, caption = "Haddock error results")
```

\section{Introduction}

much of below is from https://apps-nefsc.fisheries.noaa.gov/nefsc/ecosystem-ecology/
or
https://www.fisheries.noaa.gov/data-tools/fisheries-economics-united-states-data-and-visualizations

- The eastern continental shelf is ecologically diverse and economically important

The Northeast United States continental shelf spans from the Outer Banks of North Carolina to the Gulf of Maine. The region covers over 250,000 km$^2$ of ocean, extending over 200 km from shore in the largest areas in New England to just 30 km off shore in the southern regions. This ecologically diverse region contains approximately 18,000 vertebrate marine species. Commercial fisheries have been an important part of local economies for centuries. In 2019, New England fisheries produced $22 billion in sales, which sustained over 200,000 jobs. Maintaining a healthy ecosystem is therefore vital to sustained ecological health and economic prosperity of the region. [@nefmc20]

- Bottom trawl survey is important for monitoring population trends

Fish stocks in this highly productive and economically important region are managed by the National Oceanic and Atmospheric Administration’s (NOAA) Northeast Fisheries Science Center (NEFSC) in Woods Hole, Massachusetts. Federal biologists assess the health and abundance of each commercial fish stock using fishery-independent bottom trawl survey data that has been collected by NOAA throughout the region since 1963 (cite survey paper). The survey uses a stratified random design where bottom trawl sampling takes place in predefined strata along the eastern continental shelf **(include image from johnson_sosebee?)**. The survey has created a rich time series data set with many uses including species-specific habitat identification, analysis of how environmental conditions influence species abundance, and estimating yearly species abundance trends to help inform stock assessments and ultimately quota limits **just listed a few uses of survey change/others?**.

The survey takes place twice each year- once in the spring and again in the fall. Since most spatial analyses and  projections of future distributions typically assume a constant survey catchability and/or availability over time, NOAA's survey design includes sampling during approximately the same 2-3 week time period in each season.

- Climate change is happening
 Due to a combination of climate change and shifts in circulation, the Northeast United States continental shelf has experienced rapid warming in recent decades, resulting in a shift in spatial distributions of many species. Since stock assessment models rely on accurate descriptions of population dynamics and contemporary patterns of spatial abundance, there is concern that rapid undocumented changes in spatial distributions of species will bias future stock assessments. The implication of this is that the bottom trawl survey is actually sampling the population during a different life cycle stage than was originally assumed, which can lead to biased stock assessments. We are therefore interested in analyzing the impact of climate change on the accuracy of future stock assessment models as measured by NOAA’s ongoing bottom-trawl survey along the East coast.

**use more info from initial proposal**

- Fish are changing spatial distribution and have altered life stages (?) because of climate change
NYE paper

- Population indexing methods may be becoming biased as a result

- Briefly describe our study to test this

To test the ability of the bottom trawl survey to track population trends under shifting environmental conditions, we construct spatial models for fish where movement depend on temperature preferences. We can then consider the impact of climate change by simulating scenarios with repeating temperature patterns and those where temperature increases on average over time. In both cases we analyze the ability of stratified random sampling to track population trends.

\section{Methods}

- Describe simulation study

We construct spatial models for Yellowtail Flounder, Atlantic Cod, and Haddock on George's Bank, where movement of each species combine static species-specific habitat preferences with dynamic temperature preferences. Model dynamics are driven by dynamic temperature gradients estimated from data to create simulated data sets for each population where the true biomass is known. Using temperature gradients that repeat each year creates data sets with predictable, repeating spatial patterns, whereas using a temperature gradient that increases on average throughout the simulation leads to spatial distributions that shift over time. We conducting stratified random sampling on our simulation output to mimic the bottom trawl survey and compare the ability of contemporary indexing methods to track population trends.

\subsection{Population Model Formulation}

-- Used MixFishSim. Describe edits made to package

We use the R package *MixFishSim* (MFS) to model our populations [@dolder2020highly]. MFS is a discrete spatiotemporal simulation tool where users can model multiple species under varying environmental conditions. The package uses a delay-difference population model with discrete processes for growth, death, and recruitment of the population. We formulate the following inputs for the MFS package to address our research question:

<font size="5">*Study Area*</font>

We obtained a shapefile for the 15 strata that comprise George's Bank to use as our modeling environment. We discritized the region into a raster with 88 rows and 144 columns. Haddock inhabit all 15 strata in the domain Cod inhabit 13 strata, and yellowtail exist in 9 strata. **include figure?**

<font size="5">*Population Dynamics and Recruitment*</font>

The time step for our models is one week. MFS uses a modified two-stage Deriso-Schnute delay difference equation that models the biomass of in each cell in our study area [@dolder2020highly]. Individual terms in the formulation account for growth of mature adults, natural and fishing mortality, and the addition of new recruits. We chose to represent recruitment in the model using a Beverton-Holt formulation. Recruitment is a function of the adult biomass that existed in the previous year and is added to the population incrementally throughout each species' predefined spawning period.

--Many parameter inputs were obtained from literature

Each species is assumed to have normally distributed temperature preferences ($~N(\mu,\sigma)$). We assume Yellowtail's preferences are $N(8.75,4.25)$, while Haddock and Cod have preferences $N(9,4)$. We chose these values by combining information in the literature with temperatures recorded in the bottom trawl survey. A full list of parameters used in our model can be seen below in Table ???.

**go through each process in model and discuss edits/removals?**

<font size="5">*Temperature and Habitat Input*</font>

The package was designed to generate theoretical habitat preferences using Gaussian Random Fields that combine with hypothetical temperature gradients to simulate an entire fishery. Since we are modeling real species on the northeast continental shelf w

We obtain estimated temperature data for the region for 2012 from FVCOM **cite**.

--Describe difference between increasing and constant temperature scenarios (images?)

-- Describe each scenario that is considered

We consider three population parameter scenarios for each of our three species- a scenario where parameters result in each population increasing over time, one where the populations are relatively constant over time, and a scenario where the parameter combination results in each population decreasing over time. Each of these three scenarios is paired with a temperature gradient that repeats as well as one that increasing roughly 5 degrees calcius over the duration of the 20 year simulation.

<font size="5">*Simulating Bottom Trawl Survey and Population Indexing*</font>

-Describe post hoc sampling process and how data is used

After each simulation is complete, we mimic the bottom trawl survey by conducting stratified random sampling in each inhabited strata twice each year. The number of the samples taken in each strata reflect true values and we sample from the same weeks that the Spring and Fall surveys take place. Most strata contain enough cells to sample a unique location in each survey. For smaller strata we are forced to repeat locations. We then use the samples in contemporary population indexing methods to track population trends. Since we know the true population values in our simulations we can compare methods by comparing the error calculated from each estimate.

--Stratified mean vs VAST with and without covariates

We compare the stratified mean estimate of abundance to estimates obtained from the Vector-Autoregressive Spatio-temporal (VAST) model. The stratified mean is a typical survey-based approach that scales individual samples to the  strata-level by considering the area of each strata, before scaling to the region-level based on the relative size of each strata. VAST is a spatio-temporal statistical framework that models both abundance (biomass) and probability of occurance (presence/absence). If desired, VAST also allows users to include covariate data to better inform the model.

>>>>

---
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
output:
  pdf_document:
    toc: no
  bookdown::pdf_document2:
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: yes
  word_document: default
  html_document: default
fontsize: 12pt
header-includes:
- \usepackage{setspace}
- \doublespacing
- \usepackage{lineno}
- \linenumbers
- \usepackage[belowskip=0pt,aboveskip=0pt]{caption}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(kableExtra)) {  
  install.packages("kableExtra")
  require(kableExtra)}
```

@@

# Data Rich but Model Resistant: An Evaluation of Data-Limited Methods to Manage Fisheries with Failed Age-based Stock Assessments

Christopher M. Legault^1^, John Wiedenmann^2^ Jonathan J. Deroba^1^, Gavin Fay^3^, Timothy J. Miller^1^, Elizabeth N. Brooks^1^, Richard J. Bell^4^, Joseph A. Langan^5^, Jamie M. Cournane^6^, Andrew W. Jones^1^, Brandon Muffley^7^

^1^National Marine Fisheries Service, Northeast Fisheries Science Center, Woods Hole, MA, USA  
^2^Department of Ecology, Evolution and Natural Resources, Rutgers, The State University of New Jersey, New Brunswick, NJ, USA  
^3^University of Massachusetts Dartmouth, School for Marine Science
and Technology, MA, USA  
^4^The Nature Conservancy, RI, USA  
^5^University of Rhode Island, Graduate School of Oceanography, Narragansett, RI, USA  
^6^New England Fishery Management Council, Newburyport, MA, USA  
^7^Mid-Atlantic Fishery Management Council, Dover, DE, USA

Corresponding author: C.M. Legault (chris.legault@noaa.gov)

@@

@@

Age-based stock assessments are sometimes rejected by review panels due to large retrospective patterns. When this occurs, data-limited approaches are often used to set catch advice, under the assumption that these simpler methods will not be impacted by the problems causing retrospective patterns in the age-based assessment. This assumption has never been formally evaluated. Closed-loop simulations were conducted where a known source of error caused a retrospective pattern in an age-based assessment. Twelve data-limited methods, an ensemble of a subset of these methods, and a statistical catch-at-age model with retrospective adjustment were all evaluated to examine their ability to prevent overfishing and rebuild overfished stocks. Overall, none of the methods evaluated performed best across the scenarios. A number of methods performed consistently poorly, resulting in frequent and intense overfishing and low stock sizes. The retrospective adjusted statistical catch-at-age assessment performed better than a number of the alternatives explored. Thus, using a data-limited approach to set catch advice will not necessarily result in better performance than relying on the age-based assessment with a retrospective adjustment.

@@

closed-loop simulation, data-limited methods, retrospective analysis, management advice

@@

## Introduction
In the U.S., age-based, integrated, fisheries stock assessment models are frequently used to estimate annual stock abundance (biomass), fishing mortality rates, and management reference points [@maunder2013review]. These models must undergo peer review, where an independent panel of experts determines whether or not results from the model are suitable as the basis for determining stock status and for setting catch advice. There are a number of model diagnostics that are used to evaluate uncertainty and stability of assessment model results, but one that is commonly used and carries substantial weight during review is the retrospective pattern. A retrospective pattern is a systematic inconsistency among a series of sequential assessment estimates of population size (or other related assessment variables), based on increasing time periods of data used in the model fitting [@mohn1999rho]. These inconsistencies in assessment estimates are indicative of one or more mismatches between model assumptions and patterns in the data used to fit the model. Large or persistent retrospective patterns indicate an instability in model results, and may therefore be the basis for a peer review panel to determine that model results are not suitable for management purposes [@punt2020reject].

Many stock assessments in the Northeast U.S. have a history of strong retrospective patterns, whereby estimates of biomass are typically revised downward and estimates of fishing mortality rate are revised upward as new data are added to the model (i.e., implying systematic overestimation of biomass and underestimation of fishing mortality; [@ices2020wkforbias]). NOAA Fisheries, the New England Fishery Management Council, the Mid-Atlantic Fishery Management Council, and the Atlantic States Marine Fisheries Commission manage these stocks, and retrospective issues remain a challenge for managers when setting catch advice and tracking stock status. This problem has been particularly acute for, but not limited to, stocks in the New England groundfish complex [@nefsc02a; @nefsc05; @nefsc08; @nefsc15a; @nefsc15b; @nefsc17; @nefsc19; @deroba2010mackerel], managed under NOAA Fisheries and the New England Council’s Northeast Multispecies (Groundfish) fishery management plan. Stock assessments exhibiting retrospective patterns can be found around the world and can be associated with a wide range of assessment approaches [@ices2020wkforbias].

The magnitude of the retrospective pattern is typically measured with a statistic called Mohn’s rho [@mohn1999rho]. Mohn's rho can be used to adjust terminal year estimates of biomass in anticipation that the retrospective pattern will persist, and some accounting for the pattern will provide a more accurate estimate. Stock assessments where the so-called rho-adjusted value is outside the 90\% confidence interval of the terminal year estimate of spawning stock biomass ($SSB$) or fishing mortality rate are classified as strong retrospective patterns. In these cases, the rho-adjusted values are used for status determination and to modify the starting population for projections used to provide catch advice [@brooks2016retroforecast].

There are many possible causes for retrospective patterns, but typically there is a temporal change in either the data or a model parameter that is not accounted for in the stock assessment model [@legault2020rose; @hurtado2014rearview; @deroba2014retro]. The strong retrospective patterns seen in the region under study have required very large magnitudes of change in order to remove the retrospective pattern.  For example, the scale of missing catch needed to be three to five times the reported catch, or natural mortality needed to increase from 0.2 to near 1.0 to reproduce observed retrospective patterns; the scales of these changes have not been deemed believable by review panels. Some approaches have been used to estimate missing catch [@vanbeveren2017catch; @perretti2020sim] and increased natural mortality [@cadigan2016ss; @rossi2019inferring]. However, identifying the correct source of the retrospective pattern is difficult and using the wrong fix can lead to poor management advice [@szuwalski2017retro]. This is clearly an area where more research is needed, but currently addressing strong retrospective patterns is challenging.

There is no formal criteria in the region for rejecting an assessment based on Mohn’s rho, but large, positive values of rho for $SSB$, especially those persisting across several assessments, have played an important role in the rejection of recent age-based assessments, including Atlantic mackerel (*Scomber scombrus*), Georges Bank Atlantic cod (*Gadus morhua*), Georges Bank yellowtail flounder (*Limanda ferruginea*), and witch flounder (*Glyptocephalus cynoglossus*) [@deroba2010mackerel; @legault2014tracgbytf; @nefsc15a; @nefsc15b]. In each of these cases, and another where the assessment rejection was not based on the retrospective pattern [black sea bass, *Centropristis striatus*, @nefsc12], the Councils have relied on a variety of data-limited approaches for setting catch advice for these stocks [@mcnamee2015mafmc; @nefsc15a; @nefsc15b; @wiedenmann2015mackerel]. These approaches have all been ad-hoc, and a recent analysis suggested that some of the data-limited approaches may not be suitable for stocks in the Northeast U.S. with a history of high exploitation rates [@wiedenman2019dlm]. In addition, large, positive retrospective patterns in $SSB$ persist for a number of other stocks in the region [@nefsc19], raising concerns that additional stocks may rely on data-limited approaches in the future.

Current practice in the region requires identification of a back-up assessment approach for all age-based assessments in case the age-based assessment is rejected during peer review. These back-up approaches are required to be simple enough that only minor review is needed so that management advice can continue to be provided for the stock. While these DLMs cannot provide stock status determinations in our study because they rely on ad hoc setting of reference points, they all can provide catch advice. Therefore, there is an immediate need to identify suitable data-limited approaches for setting catch advice for stocks with age-based assessments that did not pass review.

We developed a closed-loop simulation [e.g., @punt2016mse; @huynh2022retro] to evaluate the suitability of alternative data-limited methods (DLMs) for setting target catches when age-based stock assessments fail. In particular, focus was placed on methods that use survey indices of abundance. The closed-loop simulation was designed to test the two most common hypothesized sources of retrospective pattern (missing catch or increases in natural mortality), and to evaluate performance of various methods relative to exploitation history and changes in fishery selectivity. Results of this factorial simulation study are summarized for quantities of interest that impact fisheries management advice. The goal of this work is to examine the hypothesis that catch advice from DLMs is more robust to under-reported catch or changes in natural mortality than from a rho-adjusted statistical catch at age model.

## Methods

<font size="5">*Overview*</font>

A closed-loop simulation was designed to approximate a process where an age-based assessment was rejected due to a retrospective pattern, requiring catch advice to be determined using a DLM. As such, the operating model (OM) used to define the "true" underlying biological and fishery dynamics was also age-based. The OM was run for an initial 50 year period of time (called the base period) that controls the historical population dynamics and fishing pressure, and allows for sufficient data to be simulated in the observation model to be used in the different DLMs. After the base period, a given management approach (i.e., DLM) was applied to set the target catch for the stock, which is then removed from the population. This process is repeated at a fixed interval for 40 years in what is called the feedback period. Multiple OMs were developed so that the performance of the DLMs could be compared among several sources of uncertainty that are especially common in the northeast U.S., but relevant more broadly. The set of OMs featured one of two possible patterns of time varying dynamics in the last 20 years of the base period, that if left misspecified as time invariant, would be sufficient to generate retrospective patterns resulting in the rejection of an age-based stock assessment, requiring transition to a DLM. The details of these dynamics, and the suite of factors explored in the closed-loop simulation, are described in sections below.

<font size="5">*Operating and Observation Models*</font>

The Woods Hole Assessment Model [WHAM, @miller2020wham; @stock2021wham] was used as the basis for the OM in the closed-loop simulations. WHAM is an R package and the general model is built using the Template Model Builder package [@kristensenetal2016TMB]. While WHAM can serve as a stock assessment model used to estimate parameters, it can also simulate the data needed for age-based stock assessments and DLMs given a range of input parameters. WHAM was used to simulate data with known properties during the base and feedback periods. Catch and index observations upon which the DLMs largely relied were simulated according to user supplied biological and fishery parameters for each scenario (see below). Catches during the feedback period were iteratively updated based on a DLM and harvest control rule that used the simulated observations to produce catch advice. Catch advice from a given combination of DLM and control rule was specified in two year blocks, a typical catch specification timeframe for New England and Mid-Atlantic Council managed fisheries. WHAM used these catches, along with the user supplied biological and fishery inputs, to have the simulated population respond to the DLM, thereby completing the closed-loop simulation aspect. A limit was placed on the maximum fishing mortality rate when the fishery attempted to remove the catch advice from the population during the feedback period. There was no implementation error in the removal of the catch advice otherwise, except when missing catch was the source of the retrospective pattern as described below.

The age-structured OM had ten ages, with the oldest age being a plus group. Maturity- and weight-at-age were time and simulation invariant and reflected values observed for groundfish in the region (Table 1). The OM simulated catch and age composition data for a single fishery with logistic selectivity (Table 1; see below). Annual, total catch observations (metric tons) were simulated as lognormal deviations from the underlying "true" catches with a coefficient of variation (CV) equal to 0.1. Fishery age composition data were assumed to follow a multinomial distribution with an effective sample size (ESS) equal to 200. Two fishery independent surveys were simulated and were intended to represent the spring and fall, coastwide bottom trawl surveys conducted in the region. Both surveys were assumed to have time invariant logistic selectivity and constant catchability. Annual survey observations were simulated as lognormal deviations from the underlying "true" survey catches with a CV of 0.3 in the spring survey and 0.4 in the fall. Survey age composition data were assumed to follow a multinomial distribution with an ESS equal to 100 in both seasons.

Annual recruitment was simulated as autoregressive, lag-1 (AR-1) deviations from an underlying Beverton-Holt stock-recruitment relationship with steepness equal to 0.74. The degree of correlation in the AR-1 process equaled 0.4 with a conditional standard deviation about this relationship equal to 0.5. Unfished recruitment was time- and simulation invariant and equaled 10-million age-1 fish. These stock-recruitment values were based on an average of groundfish parameters estimated for the region.

<font size="5">*Data-Limited Methods Explored*</font>

The range of DLMs evaluated was generally constrained to those that have been used or were considered plausible (e.g., based on data requirements) for the Northeast Shelf. Ultimately, thirteen DLMs were selected for evaluation. Although catch-curve analyses are not currently applied in the region, they were included here since age information is available for most of the stocks, and because @wiedenman2019dlm showed they performed well in application to groundfish stocks. Two additional DLMs (Islope and Itarget) not currently used in the region were also evaluated, as these have been tested in other applications and shown promise [@geromont2015complex; @geromont2015datapoor; @carruthers2016simpleMPs; @wiedenman2019dlm]. An ensemble of models was also considered based on recent findings that improved performance can result from combining the results from multiple models [@anderson2017superensemble; @rosenberg2018ensemble; @spence2018combineecomodels; @stewart2018ensemble]. The catch advice from the ensemble approach equaled the median of the catch advice resulting from the range of methods included in the ensemble (Table 2). This assumes an equal weighting of ensemble members. The DynLin approach was excluded from the ensemble due to the relatively long computing time required. Other methods were excluded (CC-FM, ES-FM, ES-Fstable) because they were slight variations of a more generic DLM (i.e., CC- and ES-) and including them all may have unduly overweighted the performance of the ensemble towards these methods. For the methods with multiple variations, the variant retained in the ensemble had superior performance than the alternatives based on preliminary results, or had already been considered for application in the region. The full range of methods included in this analysis were detailed below with equations (Table 2). Each method was applied to data that would lead to retrospective patterns in an age-based stock assessment and performance was evaluated using a range of metrics (see below).

Each of the methods evaluated produces a single target catch value that was fixed over a two year interval. If the methods were being applied in year $y$, then target catches are set for years $y+1$ and $y+2$ (denoted $C_{targ, y+1:y+2}$). In practice, the timing of setting target catches in the region generally occurs in late summer or early fall in between the spring and fall surveys, and before complete catch data are available. Therefore, in year $y$ complete catch data are available through year $y-1$, and survey data are available for the spring survey through year $y$ and for the fall survey through year $y-1$. Applications of DLMs in this region have used an average of the spring index in year $y$ $(I_{spr,y})$ and the fall index in year $y-1$ $(I_{fall,y-1})$ to reflect average abundance at the start of year $y$ $(\bar{I}_{y})$. For this study, the same 1 year lag was implemented for methods that use the average of both simulated indices to generate catch advice:

$\bar{I}_{y} = \frac{I_{fall,y-1} + I_{spr,y}}{2}$.

<font size="5">*Control Rules*</font>

Most DLMs do not have the ability to estimate a biomass reference point (e.g., $B_{MSY}$), which made consideration of so called biomass-based harvest control rules that reduce $F$ or catch in response to estimated changes in relative stock status impossible. Although reference points can be created for DLMs, they typically rely on local expert judgment [@harford2021harvest] and are geared towards either keeping the stock about where it is or else increasing it towards a relative amount that was thought to be good. Neither of these provide a proxy for maximum sustainable yield reference points, but might instead provide pretty good yield [@hilborn2010pgy].

Lack of clarity exists, however, on whether the catch advice from DLMs should be used directly or reduced to account for uncertainty. In the U.S. management system, an overfishing limit is the catch that would result from applying $F_{MSY}$, whereas an acceptable biological catch is a catch reduced from the overfishing limit to account for scientific uncertainty. Each DLM was evaluated using two harvest control rules: 1) the catch advice from a given DLM was applied directly and assumed to serve as a proxy for the catch associated with $F_{MSY}$ (catch multiplier = 1), and 2) the catch advice from a given DLM was reduced by 25\% to account for unspecified scientific uncertainty (catch multiplier = 0.75). The case where catches were reduced by 25\% was intended to reflect a common default control rule in the region that uses 0.75$F_{MSY}$.

<font size="5">*Application of a Statistical Catch-at-Age Assessment (SCAA)*</font>

A SCAA model was also applied to all scenarios to generate catch advice for comparison with the DLMs. Although virtual population analysis (VPA) is also used for some age-based assessments in the region, SCAA models are more widely used. Applications of the SCAA model assumed that the assessment had the correct underlying structure for selectivity, and CVs and ESS were specified at their true underlying values. The SCAA model estimated annual recruitment deviations assuming no underlying stock-recruit relationship, annual fully-selected fishing mortality rates, fishery and survey selectivity parameters (logistic), abundance-at-age in year one of the period being assessed, and survey catchabilies. Mohn’s rho was calculated (7 year peels) for abundance at age for all model fits during the feedback period and used to retro-adjust abundance at age for projections (divided by one plus Mohn’s rho; [@brooks2016retroforecast]). Catch advice was determined by specifying fully-selected $F = 0.75F_{40\%}$, always assuming M=0.2. All life history parameters were fixed at their correct value, except for the natural mortality rate when it was the source of the retrospective pattern.

<font size="5">*Study Design*</font>

In addition to the two control rules applied for each DLM described above, three aspects of the OM were varied in a full factorial study design: fishing history, fishery selectivity, and cause of the retrospective pattern (Table 3). Two variants of fishing history were considered, with fully selected fishing mortality during the base period either constant at a level equal to $2.5F_{MSY}$ (always overfishing) or equaling $2.5F_{MSY}$ in the first half of the base period then a knife-edged decline to $F_{MSY}$ for the second half of the base period. These patterns in fishing mortality rate were based on observed patterns for Northeast groundfish [@wiedenman2019dlm]. These two different fishing intensities during the latter half of the base period led to different starting conditions for the feedback period.

Two variations of the OM were considered with either time invariant, asymptotic, fishery selectivity in the base and feedback periods, or a change in selectivity after the first half of the base period so that the age at 50\% selectivity increased from approximately 3.7 to 5 (Table 1). The asymptotic selectivity pattern was based on Northeast groundfish fishery selectivity patterns. The change in the selectivity pattern when selectivity varied through time approximated an increase in mesh size in the fishery to avoid younger fish.

Two different sources of stock assessment misspecification leading to retrospective patterns were considered, temporal changes in natural mortality and misreported catch. The degree to which natural mortality and unreported catch changed through time was determined by attempting to achieve an average Mohn’s rho of approximately 0.5 for $SSB$ when an SCAA model (i.e., configured using WHAM) was used to fit the simulated data. We also fit the same SCAA configuration to data without misspecified M or catch to verify that retrospective patterns were not present on average (see Supplemental Materials Figure S1). A third source of misspecification was also attempted, time varying survey catchability, but this source of misspecification was unable to produce severe enough retrospective patterns and was abandoned.

For the natural mortality misspecification, the true natural mortality changed from 0.2 to 0.32 in scenarios where the fishing history was always overfishing or from 0.2 to 0.36 when the fishing history included a reduction from overfished to $F_{MSY}$, with the differences between fishing histories necessary to produce the desired retrospective pattern severity (see Supplemental Materials Figures S2 and S3). In each case, natural mortality trended linearly from 0.2 to the higher value between years 31 and 40 of the base period and held constant at the higher level for years 41-50. Natural mortality remained constant at the higher level throughout the feedback period. Those DLMs that required natural mortality as an input parameter used the value from before any change in natural mortality (0.2) because the change in natural mortality is meant to be unknown.

For catch misspecification, a scalar multiple of the true catch observation is provided as the observed catch to the DLMs. The scalar is 0.2 when fishing intensity was always overfishing and for both selectivity patterns, 0.44 when the fishing history included a reduction to $F_{MSY}$ and with time variant selectivity, or 0.40 when the fishing history included a reduction to $F_{MSY}$ and selectivity was time invariant. The shift in scalar trended linearly from 1 to the lower value between years 31 and 40 of the base period and remained at the lower value for years 41-50. These scalars were applied only to the aggregate catch so that they affect all catches at age equally. When catch misspecification was applied in conjunction with a DLM during the feedback period, the true catch in the OM equaled the catch advice provided by the DLM multiplied by the inverse of the scalar multipliers (i.e., the true catches were higher than the DLM catch advice). Thus, when the scalar multipliers were applied to the true catch from the OM in order to provide observed catches at the next application of the DLM, the observed catch equaled the catch advice from the previous application of the DLM, on average. In other words, managers and analysts would be given the perception that the DLM catch advice was being caught by the fishery, when in fact the true catches were always higher. This meant that the source of the retrospective pattern continued in the feedback period. The magnitude of the retrospective pattern in the feedback period varied due to the observation error applied in each realization (See Supplemental Materials Figure S4).

Fourteen methods for setting catches were explored (13 DLMs and the SCAA) and were applied to all 16 scenarios, which created 224 factorial combinations in the study design. For each element of the full factorial combinations, 1,000 simulations were conducted. The simulations used the same random number seeds across all combinations in the study design resulting in the same patterns of recruitment deviations and observation errors. Two DLMs (AIM and ES-Fstable) had two failed simulations each, which were caused by relatively high catch advice (i.e., requiring relatively high $F$) that triggered errors in the Newton-Raphson iterations used to determine the $F$ that would produce the desired catch. This small number of failures was unlikely to effect results and conclusions, and so were not considered further.

<font size="5">*Performance Metrics*</font>

Six metrics thought to be of broad interest were reported here, each calculated and reported separately for a short-term (i.e., first six years of the feedback period) and long-term (i.e., last 20 years of the feedback period) period. These metrics were selected to represent the tradeoffs in terms of benefits to the fishery and risks to the stock. The specific metrics reported were: $\frac{SSB}{SSB_{MSY}}$, $\frac{F}{F_{MSY}}$, catch relative to $MSY$, interannual variation in catch [@amar2010mse], number of years of overfishing ($F > F_{MSY}$), and number of years of the stock being overfished ($SSB < 0.5 SSB_{MSY}$).

## Results

Overall performance varied widely across methods, and the individual performance of a method was sensitive to the different scenarios explored. Performance for each method was sensitive to the source of the retrospective pattern (missing catch or M), the exploitation history, when in the feedback period the metric was calculated (short- or long-term), and whether or not a 25% buffer was applied when setting the catch advice from a given method. Overall, similar results occurred for the scenarios with one or two selectivity blocks, so the impact of the selectivity scenarios was not discussed further.

<font size="5">*Aggregate performance*</font>

In Figure \ref{fig:PMsallscenarios}, the inner quartiles and medians for all performance measures are shown, calculated across all scenarios combined. In general, methods that resulted in high mean $F / F_{MSY}$ (Figure \ref{fig:PMsallscenarios}B) resulted in lower stock biomass (Figure \ref{fig:PMsallscenarios}A), more years of overfishing (Figure \ref{fig:PMsallscenarios}E) and of being overfished (Figure \ref{fig:PMsallscenarios}F), and vice-versa. Higher $F$ values were also associated with higher catches (Figure \ref{fig:PMsallscenarios}C), on average, and a greater variability in catch, but there were some methods that produced lower $F$ values that also resulted in high catch variability (CC-FM, CC-FSPR; Figure \ref{fig:PMsallscenarios}D).

A number of methods performed poorly overall, resulting in high exploitation rates and low stock size, on average (Figure \ref{fig:PMsallscenarios}). These methods include AIM, three of the four expanded survey biomass methods (ES-FM, ES-FSPR, and ES-Fstable), and the Skate method. The Itarget and ensemble methods also resulted in $SSB < SSM_{MSY}$ and $F > F_{MSY}$, on average, though departures from the MSY levels were not as severe as the other methods (Figure \ref{fig:PMsallscenarios}). The remaining methods (CC-FM, CC-FSPR, DynLin, ES-Frecent, Islope, Ismooth, and SCAA) were able to limit overfishing and keep biomass above $SSB_{MSY}$, on average, although for four of these methods (CC-FM, CC-FSPR, DynLin, and Ismooth) biomass was more than 50% higher than $SSB_{MSY}$ (Figure \ref{fig:PMsallscenarios}). Principal components analysis of the median values for all methods and metrics resulted in groupings similar to those noted above (see Supplemental Materials Figure S5).

<font size="5">*Scenario-dependent performance*</font>

The source of the retrospective pattern had a large impact on results for a given method. The relationship between $SSB/SSB_{MSY}$ and $C/MSY$ is shown across scenarios for the different sources of retrospective error. Stock size and catch (relative to MSY levels) are clustered for many of the methods with no overlap between $M$ and unreported catch sources (AIM, ES-FM, ES-FSPR, ES-Fstable, Itarget, Skate, Ensemble, and SCAA). For all of these methods, $SSB/SSB_{MSY}$ was lower when unreported catch was the source of the retrospective pattern, and $C/MSY$ was also lower except for the Itarget and the SCAA methods compared to the scenarios when increased natural mortality was the source of the retrospective pattern (Figure \ref{fig:SSmsyvsCMSY}). The source of the retrospective pattern also had a large impact on the other performance measures (Figure \ref{fig:PMsbyretroF}). In general, when unreported catch was the source of the retrospective pattern, interannual variability in catch was higher, overfishing was more frequent and with a larger $F/F_{MSY}$, and the stock had a higher risk of being overfished compared to the scenarios when increased natural mortality was the source of the retrospective pattern (Figure \ref{fig:PMsbyretroF}). Six methods (AIM, ES-FM, ES-FSPR, ES-Fstable, Itarget, Skate, Ensemble) resulted in overfishing in nearly every year of the feedback period (often with very high $F/F_{MSY}$) when missing catch was the source of the retrospective pattern (Figure \ref{fig:PMsbyretroF}B, \ref{fig:PMsbyretroF}E). In contrast, all methods except Skate, AIM, and ES-Fstable had low $F/F_{MSY}$, high $SSB/SSB_{MSY}$, and few years of being overfished when increased natural mortality was the source of the retrospective pattern (Figure \ref{fig:PMsbyretroF}B, \ref{fig:PMsbyretroF}A, \ref{fig:PMsbyretroF}F). The $C/MSY$ when increased natural mortality was the source of the retrospective pattern varied widely with some DLMs well below 1.0 and others well above (Figure \ref{fig:PMsbyretroF}C). The SCAA method also resulted in frequent overfishing in the missing catch scenario, but less so when the stock was more depleted at the start of the feedback period (Figure \ref{fig:PMsbyretroF}F).

Exploitation history also impacted the performance of many of the other methods. For four methods (Islope, Ismooth, DynLin and ES-Frecent), exploitation rates were higher when the stock experienced overfishing for the entire base period, but the impact was more dramatic in the short-term. Over time as these methods were used, $F$ declined and remained below  $F_{MSY}$ in the long-term (Figure \ref{fig:FFmsybyscenario}A), allowing stock recovery. The majority of the other methods also resulted in greater exploitation rates in the short-term, though some methods kept $F/F_{MSY} < 1$ regardless of the time-period (CC-FM, CC-FSPR, and SCAA), while others (AIM, ES-Fstable, Skate, Ensemble) kept $F/F_{MSY} > 1$ over the short- and long-term (Figure \ref{fig:FFmsybyscenario}A). For the ES-FM and ES-FSPR methods, there was not a consistent pattern in exploitation rates when comparing the short- and long-term periods (Figure \ref{fig:FFmsybyscenario}A).
